---
title: "Physiology"
output: html_document
date: "2023-09-26"
---

This script analyzes and plots data for Developmental Timeseries 2023 physiology measurements including protein, carbohydrates, and lipids, and correlations between these variables. This script is based off of AH's [script](https://github.com/AHuffmyer/SymbioticIntegration/blob/main/Mcap2021/Scripts/Physiology.Rmd).

# Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)

getwd()
```

# Set plot theme 

```{r}
custom_theme<-theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        legend.text=element_text(size=12, color="black"), 
        legend.title=element_text(size=12, face="bold", color="black"))
```

# Protein

## Load data 

Read in datafiles

```{r}
protein <- read.csv("2023/data/Physiology/SolubleProtein_Data.csv")
prot_meta <- read.csv("2023/data/Physiology/SolubleProtein_Meta.csv")
meta <- read.csv("2023/data/Tube_Meta.csv")
```

## Prepare dataframes 

Remove unneeded rows from meta dataframe 

```{r}
meta <- meta %>%
  filter(Metric == "Physiology")
```

Merge Files and rename columns

```{r}
protein <- merge(prot_meta, protein, by = c("Well", "Run"))
protein <- left_join(protein, meta, by=c("TubeID"))

# Blank correction for each run separately

Blank <- protein %>% 
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(x562))

protein$abs.corr <- protein$x562 - Blank$blk.avg
```

Plot regressions for each run.  

```{r}
protein %>% 
  filter(Sample.Type == "Standard")%>%

  ggplot(aes(x=Concentration, y=abs.corr))+
  facet_wrap(~ Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  ggtitle("Run")+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = after_stat(eq.label))) +
  custom_theme
```

Run a loop to extract regression for each run and calculate protein concentration.  

```{r}
run_list<-c("1", "2", "3", "4", "5", "6")

df_protein = data.frame()

for (run in 1) {

  #subset data
Standard <- protein %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==run)
  
#generate standard curve equation
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 

#select samples
Samples <- protein %>% #subsetting Samples
  filter(Sample.Type == "Sample ")
  filter(Run==run)

#calculate concentration
Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration

#add run column
#Samples$Run<-run

#normalize to homogenate volume
#Samples$Protein.ug <- Samples$Concentration * (Samples$Resuspension_volume / 1000) * (25/Samples$Homo_vol)

#join dataframes
df <- data.frame(Samples)
df_protein <- rbind(df_protein,df)

} 
```


Plot results 

```{r}
# Order the 'hpf' variable in a specific order
df_protein$hpf <- factor(df_protein$hpf, levels = c("1 hpf", "4 hpf", "9 hpf", "14 hpf", "22 hpf", "28 hpf", "48 hpf", "72 hpf"))

Protein.Plot <- df_protein%>%
  ggplot(aes(x=hpf, y=Concentration, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  scale_fill_manual(values = c("blue", "red", "gray")) +
  custom_theme; Protein.Plot
```

## Calculate Ratio of host:holobiont   

Calculate the ratio of host:holobiont protein. 
```{r}
ratios<-df_protein%>%
  dplyr::select(TubeID, Treatment, Fraction, Concentration, hpf)%>%
  group_by(TubeID, Fraction, Treatment, hpf)%>%
  dplyr::summarise(mean_protein=mean(Concentration))%>%
  spread(Fraction, mean_protein)%>%
  mutate(ratio=Host/Holobiont)%>%
  mutate(sholo_ratio=Symbiont/Holobiont)
  
```

Plot Host:Holobiont ratios  

```{r}
Ratio.Plot <- ratios%>%
  ggplot(aes(x=Treatment, y=ratio, colour=Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
  scale_colour_manual(values=c("blue", "red"))+
  scale_fill_manual(values=c("blue", "red"))+
  xlab("Temperature") + 
  ylab("Host:Holobiont Protein") + #Axis titles
  custom_theme; Ratio.Plot 
```

Test for differences with an anova.  

```{r}
model<-aov(ratio~Treatment*hpf, data=ratios)
summary(model)
```




